---
title: "homework2"
author: "ht2611"
date: "2022-10-05"
output: github_document
---

### Problem 0

```{r load_libraries}
library(tidyverse)
library(readxl)
```

### Problem 1

update `entry` from `yes` / `no` to a logical variable. `Route` columns 8-11 should be character for consistency with 1-7.

```{r}
trans_ent = 
  read_csv(
    "data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>% 
  janitor::clean_names() %>% 
  select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, exit_only, vending, entrance_type, 
    ada) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

these data are not "tidy": route number should be a variable, as should route. That is, to obtain a tidy dataset we would need to convert `route` variables from wide to long format. 

Tuses `distinct()` to obtain all unique combinations. 

```{r}
trans_ent %>% 
  select(station_name, line) %>% 
  distinct
```

filters according to ADA compliance as an initial step. T

```{r}
trans_ent %>% 
  filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

exclude station entrances that do not allow vending. Then, we focus on the `entry` variable -- this logical, so taking the mean will produce the desired proportion 

```{r}
trans_ent %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```

convert `route` from wide to long format. filtering to focus on the A train, and on ADA compliance; selecting and using `distinct` 

```{r}
trans_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct

trans_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```



### Problem 2

```{r}
data1=read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",sheet="Mr. Trash Wheel",skip=1)%>%
  janitor::clean_names() %>%
  select(1:14)%>%
  na.omit()%>%
  mutate(sports_balls=as.integer(round(sports_balls)),type="Mr.")
```

```{r}
data2=read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",sheet="Professor Trash Wheel",skip=1)%>%
  janitor::clean_names() %>%
  select(1:14)%>%
  na.omit()%>%
  mutate(dumpster=as.character(dumpster),sports_balls=as.integer(round(sports_balls)),type="Professor")
final_data=full_join(data1,data2)

```
All the data have 15 variables. The data1 has 453 observations. Data2 has 71 observations. Final data has 524 observations. The variables such as "weight" and "volume". The total weight is 135.5. The total number is 856.
### Problem 3
```{r}
data3=read_csv("./data/pols-month.csv")%>%
  janitor::clean_names() %>%
  separate(mon,into=c("year","month","day"),sep="-")%>%
  mutate(
    month=recode(month,"01"="Jan","02"="Feb","03"="Mar","04"="Apr",
                 "05"="May","06"="Jun","07"="Jul","08"="Agu","09"="Sep","10"="Oct","11"="Nov","12"="Dec"
               ),
    president=ifelse(prez_gop==0,'dem','gop'))%>%
  select(-prez_gop,-prez_dem,-day)
data3$year=as.numeric(data3$year)

```
```{r}
data4=read_csv("./data/snp.csv")%>%
  janitor::clean_names() %>%
  separate(date,into=c("month","day","year"),sep="/")
data4$year=as.numeric(data4$year)
data4$month=as.numeric(data4$month)
data5=data4%>%
  mutate(data4,year=ifelse(year<=20,year+2000,year+1900),
         month=recode(month,"01"="Jan","02"="Feb","03"="Mar","04"="Apr",
                 "05"="May","06"="Jun","07"="Jul","08"="Agu","09"="Sep","10"="Oct","11"="Nov","12"="Dec"
               ))%>%
  select(year,month,close)
```
```{r}
data6=read_csv("./data/unemployment.csv")
data6=pivot_longer(data6,Jan:Dec,names_to="month",values_to = "percentage unemployment")%>%
  janitor::clean_names()
final_data2=full_join(data3, data5)
final_data3=full_join(final_data2,data6)
summary(final_data3)
```
Data3 has 822 observations of 9 variables, it has variables such as gov_gop and president, Data5 has 787 observations of 3 variables. Data6 has 816 observations of 3 variables. Data5 and Data6 have the same variables. The final_data3 has 1639 observations of 11 variables. The range of year is 68. The key variables are year, month.